Locality and Computational Reliability of Linear Response Calculations for Molecular Systems

### 2018-10-24 Response_referees.tex

\documentclass[11pt,a4paper]{article}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{amssymb,amsmath}
\usepackage[english]{babel}

% Parametri di stampa
\setlength{\topmargin}{-2.5cm} \setlength{\oddsidemargin}{0.3cm}
\setlength{\evensidemargin}{0.3cm}
%
\textheight=23.0truecm \textwidth=16.0truecm
%
\headheight=1.0cm \headsep=1.0cm
%
\renewcommand{\baselinestretch}{1.1}
\setlength{\unitlength}{1mm}

\parindent=6pt

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\al}{\alpha}
\newcommand{\br}{\bar{\rho}}
\newcommand{\bg}{\bar{g}}
\newcommand{\bpi}{\bar{\pi}}
\newcommand{\bra}{\langle}
\newcommand{\ket}{\rangle}
\newcommand{\sst}{\scriptscriptstyle}

\newcommand{\op}[1]{\hat {#1}}

\def\sst{\scriptscriptstyle}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Response to the first referee}

We wish to thank the referee for the positive review on the paper. Here we report a list of the modification performed in the text according
to his/her suggestions:
\begin{enumerate}
  \item \emph{point concerning the dimension of the simulation box needed to achieve the convergence of the delocalized states of the $CO$ molecule}.

  Clarify that delocalized (fluctuation) states never achieve convergence (continumm collapse). What is expected to converge is the spectral function, or declined in the present case the real and imaginary part of the dynamical polarizability. To give a reasonable value from the literature we can make usage of the results of achieved by Baroni for the benzene molecule... 
  \item \emph{point concerning the convergence rate of the delocalized vs. localized states (which one? fluctuation or excitation?) with respect
  to the grid spacing}

 Actually I don't know but I do not expected relevant differences, since localized and delocalized states differ in their asymptotic properties whereas the grid spacing convergence is influenced by the fast oscillation of the functions.

 To present a more detailed answer we could perform a convergence analysis w.r.t the gridspacing looking at the convergence rate of the energies of say 16 virtual orbitals of the $CO$ molecule. In this way we can see if the rate of the bound-state ones differs from the others.

 Lastly, the values of the grid spacing used for the calculations have been added in appendix C.

\item Answer to the other minor points...

\end{enumerate}

\section*{Response to the second referee}

We thank the referee for its comment and criticism of the paper.
We try to answer to the critical points raised out by the review:

\begin{enumerate}
 \item \emph{on the relevance of the symmetry of the molecule in determining the delocalized vs localized character of the an excitation}

 In the paper we have introduced a the dyadic representation of excitation $\op E_a$ operator (eq. 16) and we have presented an argument to
 demonstrate that the associated excited states $\phi_p^a,\chi_p^a$ are localized only if the condition $\Omega_a<\epsilon_p$ is satisfied. 
 The set of $p$ labels relevant for the construction of a specific excitation depends of the symmetry...

 %The structure of eq. 16 states that when both localized and delocalized excited states partecipate to the excitation

 \item \emph{regarding the localized excitations which can still be found above the IP threshold (Fig 5a), is there any scope to target these, to the exclusion of other states so that a set of just the localizedexcitations can be calculated at reduced computational cost compared to the whole set?}

 Answer using the results for the statical polarizability described in section III.C...

 \item \emph{exclusion ''by design'' of the delocalized part of the excitation spectrum}

 On the base of what written by the referee it seems to me that this sentence could be interpreted in two different way.

 The first one concerns with the removal of the contribution of the delocalized states in a specific excitation. This

 The second one indicate what happens when one uses a basis able to express localized states only...

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-11-05 AnalyticStructure_main.tex
update

\section{Introduction}
\added[id=dal]{Testo nuovo}.
\deleted[id=dal]{lallero}

Since the days of their foundations, disciplines like Computational Physics or Quantum Chemistry always had to deal with the problem of the 
computational reliability of the results. More specifically, a computational result may be considered reliable for various reasons.
For example, it may be in agreement with some previous observation or measurement that the employed computational model is susceptible to reproduce.
The model itself may in this way be classified as ``accurate''; it can then be employed to extract, \textit{in silico},
other quantities that might be then used to \emph{predict} or \emph{interpret} experiments.
Another concept of reliability is the concept of  ``precision'', which is more focused to the actual results
that a given computational model provides for a specific set of quantities, regardless of their potential agreement with 
experimental data. The aim of precise computational approach is to reduce the \emph{computational} uncertainties
of quantities extracted out of a well-defined model, and provide reference results, to which other computer codes
employing the same model can be compared.

A recent, remarkable example of precision-driven initiative is provided by the DeltaCodes project~\cite{deltaTest2016},
where a grassroot community of code-owners struggled to extract well-defined computational results in the context of 
Density Functional Theory (DFT) calculations. Ground-State quantum mechanical quantities as lattice constants and bulk modulus of 
elementary crystals extracted from different codes were compared with each other, with the objective of reducing the uncertainties 
on such data, and in some sense, to define a \emph{calibration} procedure for a Solid-State Physics DFT code. 
Any other code able to extract the same quantities with the same approximations may be compared with the 
results of the community in order to assess its computational reliability.

For a theoretical approach that is based on numerical calculations, for which no analytic reference solution exists, 
reducing the computational uncertainty is the only possible way to shed light on the predictivity power of the
model. In other terms, the ``accuracy'' of a result with respect to experimental data may be reliably quantified
only when the computational uncertainty is guaranteed to be significantly lower than the found discrepancy.
For communities like DFT, such a ``calibration'' of computer codes among each other is a discipline that is still in its infancy;
the DeltaCodes initiative and the foundations of DFT are separated in time by more than 50 years, and the computational
model employed in the comparison (the Perdew-Burke-Erzerhof functional~\cite{PBE}) has been introduced more than 20 years ago.

@@

In this paper, we focus our attention to Linear-Response treatments of \emph{molecular} systems, 
in particular referring to LR-TDDFT calculation. We will investigate the impact that 
\emph{open} (i.e. Isolated) Boundary Conditions (BC) that have to be imposed to the system will have on the computational evaluation of LR quantities.
In order to do this, we perform analytic considerations on the equations beyond LR treatment.
Such considerations will enable us to distinguish quantities that \textit{a priori} can be compared among different treatments - and therefore tested for reproducibility - 
from others that would \emph{explicitly} depend of the employed computational treatment. 
In particular, we will discuss the interplay between reproducible LR quantities and the \emph{locality} of the related objects.

@@

Let us start our discussion by revisiting the impact that the Isolated BC have on the 
solutions of the time-independent Schr\"odinger Equation.
We here present well-known concepts of quantum mechanics, under a perspective that
will turn out to be very useful in the forthcoming considerations about linear-response quantities.

@@

We can therefore introduce an explicit representation of the 
response density parametrized as
\be\lb{rhoPrimeFluctuationStateDef1}
\dm'_\Phi(\omega) = \sum_p\left(\ketbra{\psi_p}{f_p^\Phi(-\omega)} + \ketbra{f_p^\Phi(\omega)}{\psi_p}\right) \;.
\ee
Here we have introduced the set of $\omega$-dependent \emph{fluctuation states} 
$\ket{f_p^\Phi(\omega)} = \dm'_\Phi(\omega) \ket{\psi_p}$, that have values in the unoccupied subspace of $\hnot$. 
An analysis of the Liouville equation \eqref{LiouvillianRhopomegaDef1} for the response density written in this fashion evidences that the equations of motion of fluctuation states (FS) are 
written as a modified Sternheimer equation~\cite{mahan1980,giustino2012,giustino2014}
\be\lb{fluctuationStateEqMotion1}
\left[\omega - (\hnot-\eps_p)\right]\ket{f_p^\Phi(\omega)} = \op Q_0(\op\Phi(\omega)+\op V'[\dm'_\Phi](\omega))\ket{\psi_p} \;,
\ee
and we observe that the first-order Hamiltonian contains, 
apart from the perturbing field, a further term $\op V'[\dm'_\Phi]$ due to the density-dependence of $\hnot$.
It is easy to see that the solution of equation \eqref{fluctuationStateEqMotion1} can be casted in the form of
a set of scattering problems:
\be\lb{rhoPrimeFluctuationStateDef2}
\ket{f_p^\Phi(\omega)} = \GH(\omega+\epsilon_p)\left(\op V\ket{f_p^\Phi(\omega)} + \ket{s_p^\Phi(\omega)} \right)\;,
\ee
with the state
$
\ket{s_p^\Phi(\omega)} = \op Q_0(\op \Phi(\omega) +\op V'[\dm'_\Phi](\omega) )\ket{\psi_p} 
$ as a source term.

It can be seen that for each choice of $p$, the value $\omega = |\eps_p|$ represents the threshold level 
that governs the localization properties of the FS.
As a \emph{stringent} hypothesis, let us now assume that the source term $\ket{s_p^\Phi(\omega)}$ is localized, \textit{i.e.} its evaluation can be
restricted on a bounded real-space domain.
Even when such locality argument holds true, for $\omega > |\eps_p|$, 
the Helmholtz kernel behaves as a spherical wave and  gives raise to a $\ket{f^\Phi_p(\omega)}$ which is \emph{delocalized} in the entire real-space domain.
Such delocalization is an intrinsic consequence of the value of $\omega$, and does not depend on the particular choice of 
$\op \Phi$.
Only for $\omega < |\eps_p|$ the kernel contains an exponential damping factor and $\ket{f^\Phi_p(\omega)}$ exhibits a bound-state behavior in the long range, 
once again provided that the above locality arguments on $\ket{s_p^\Phi(\omega)}$ -- hence on $\op V'$ and $\Phi$ -- are valid.

Such locality property of the fluctuation state can be directly confirmed for static perturbations at $\omega=0$.
Let us consider a static perturbing operator $\op\Phi_s$ that can be restricted to a finite domain. 
Any local operator, like \emph{e.g.} a static electric field $\mathbf F$, with associated scalar potential $\Phi_s[\mathbf F](\r) = -\mathbf F \cdot \r$, would satisfy such property.
We consider the perturbed Hamiltonian $\hnot[\dm_{\Phi_s}] + \op\Phi_s[\mathbf F]$, where $\dm_{\Phi_s}$ represents the GS density of the perturbed system, and assume that the 
perturbation strength ($|\mathbf F|$ in the above example) is sufficiently small to be consistent with a LR treatment of the problem. 
In this case, the occupied eigenfunction $\ket{\psi_p^{\Phi_s}}$ of the perturbed Hamiltonian can be extracted with traditional GS techniques, 
as from the discussion around Eq.~\eqref{SEHelmholtz1}, they exhibit bound-state behavior.
\begin{figure}[t]
\includegraphics[scale=0.68]{Fig1_CO_statPolvsBox.eps}
\caption{\label{co_alphaStatic}(Color online) Convergence of the statical polarizability $\alpha_{zz}$ (blue continuous line) and of the ground state dipole $d_z$ (red dotted line) 
of $CO$ molecule as a function of the size of the computational domain. The inset evidences the relative convergence (in percent) of the two quantities w.r.t. the corresponding value associated to the largest 
box.}
\end{figure}

@@

\subsection{Strong and weak convergence}

@@

In view of reproducibility, the scalar product of Eq.~\eqref{LinearResponseFunctDef1}, defining the observable fluctuation, may be compared among various numerical treatments.
However, even though directly related to experimental results, such quantity is not by itself a property of the molecule, as it is a function of $\omega$ that explicitly depends 
on the perturbation operator $\op \Phi$, not to mention the target observable $\op O$. Alternatively, we may consider only the fluctuation state, but it is still a $\op \Phi$-dependent wavefunction, 
with a potential local behavior only for $\omega$ below the ionization threshold. Hence, comparing fluctuation states among diverse numerical treatments would be a 
challenging task.

It is therefore much more interesting to concentrate on LR quantities that are defined independently from the perturbing operators and target observables.
This can be achieved by studying the 
\emph{susceptibility functional}, defined implicitly trough the operator equation  
\be\lb{SusceptibilityDef1}
\dm'(\omega) = \left(\omega -\mathcal L \right)^{-1} \commutator{\op \Phi(\omega)}{\dmnot} = \sop{{\cal \chi}}(\omega) \op\Phi(\omega) \;. 
\ee
The susceptibility (super)operator $\cal \chi(\omega)$ is therefore closely related to the resolvent of the Liouvillian of Eq.~\eqref{LiouvillianRhopomegaDef1}, which is completely defined 
from the molecule Hamiltonian $\hnot[\dm]$.
The ``excitation modes''  of the molecule are defined through the \emph{excitation operators}, satisfying
\be\lb{ExcitationOperatorsDef1}
\Liouv \op E_a = \Omega_a \op E_a \qq \op{\tilde E}_a \Liouv = \Omega_a \op{\tilde E}_a \;,
\ee
together with the operator orthonormalization condition %$\trace{\op{\tilde E}\op E'} = \delta(E-E')$. 
\be\lb{orthoExcitatioOpDef1}
\trace{\op{\tilde E}_a\op E_b} = \delta_{ab} \;.
\ee
Excitations defined in that way may be considered as a basis of the operators that satisfy the transverse condition as in  
Eq.~\eqref{RhopTransverseDef1}, and enable us to express the susceptibility as a spectral decomposition:
\be\lb{RhopExcitationDef1}
\sop{{\cal \chi}}(\omega) \cdot   = \sum_{\{a\}} \op E_{a} \,
\frac{\trace{\commutator{\op{\tilde E}_a}{\dmnot}\cdot}}{\omega-\Omega_a} \;.
\ee

\subsection{The analytic structure of linear susceptibility}
Each excitation ``mode'' of the system $\op E_a$ with energy $\Omega_a$, can be associated to a set of states $\{\phi^a_p,\chi^a_p\}_{p \in a}$,
entirely defined in the unoccupied subspace. 
In this way we can parametrize, without loss of generality, the excitation operators of the system as:
\begin{align}\lb{ExcitationOperatorsDef2}
\op E_a &= \sum_p \left( \ket{\phi^a_p}\bra{\psi_p} + \ket{\psi_p} \bra{\chi^a_p}\right)\;, \nn \\
\op{\tilde E}_a = \commutator{\dmnot}{\op E^t_a} &= \sum_p \left( \ket{\psi_p}\bra{\phi^a_p} - \ket{\chi^a_p} \bra{\psi_p}\right)\;.
\end{align}
For each choice of $\{\phi^a_p,\chi^a_p\}_{p \in a}$, these objects represent, respectively, the state in which $\ket{\psi_p}$
is excited -- or from whom it decays -- when the system is subject to the \emph{monochromatic} perturbation $\op \Phi_a \equiv \commutator{\dmnot}{\op E_a}$.
We denote with the shorthand $p \in a$ the fact that the sum of Eq.~\eqref{ExcitationOperatorsDef2}
is in general restricted only to a $a$-dependent subset of the occupied orbitals, depending on the particular symmetry 
of the excitation.
Excited states  satisfy the normalization condition
\be\lb{ExcitedStateOrthNormDef1}
\sum_p \left(\brket{\phi_p^a}{\phi_p^b} - \brket{\chi_p^b}{\chi_p^a}\right) = \delta_{ab} \;. 
\ee
By applying \eqref{ExcitationOperatorsDef1} to the above parametrization
of the excitations, we obtain the following Sternheimer-like equations for the excited states:
\begin{align}\lb{ExcitationOperatorsDef3}
&\left[\Omega_a-(\hnot - \eps_p)\right] \ket{\phi_p^a} = \op Q_0 \op V'[\op E_a] \ket{\psi_p} \;, \nn\\
&\bra{\chi_p^a}\left[-\Omega_a-(\hnot - \eps_p)\right] = \bra{\psi_p} \op V'[\op E_a] \op Q_0  \;.
\end{align}
By writing these equations in a basis set of virtual states we obtain the well-known Casida's eigenvalue equation\cite{CasidaBook}
associated to the excitation energy $\Omega_a$ (see appendix \ref{casida}).
The spectrum is symmetric w.r.t the inversion of the eigenvalues 
$\Omega_a \rightarrow -\Omega_a$ and the associated excitation with negative energy are described by the transposed pair $\{\chi^a_p,\phi^a_p\}$.
We concentrate our analysis on the sector of positive energies, i.e. $\Omega_a > 0$.

>>>>

\section{Introduction}

% Since the days of their foundations, disciplines like Computational Physics or Quantum Chemistry always had to deal with the problem of the
% computational reliability of the results. More specifically, a computational result may be considered reliable for various reasons.
% For example, it may be in agreement with some previous observation or measurement that the employed computational model is susceptible to reproduce.
% The model itself may in this way be classified as ``accurate''; it can then be employed to extract, \textit{in silico},
% other quantities that might be then used to \emph{predict} or \emph{interpret} experiments.
% Another concept of reliability is the concept of  ``precision'', which is more focused to the actual results
% that a given computational model provides for a specific set of quantities, regardless of their potential agreement with
% experimental data. The aim of precise computational approach is to reduce the \emph{computational} uncertainties
% of quantities extracted out of a well-defined model, and provide reference results, to which other computer codes
% employing the same model can be compared.
% 
% A recent, remarkable example of precision-driven initiative is provided by the DeltaCodes project~\cite{deltaTest2016},
% where a grassroot community of code-owners struggled to extract well-defined computational results in the context of
% Density Functional Theory (DFT) calculations. Ground-State quantum mechanical quantities as lattice constants and bulk modulus of
% elementary crystals extracted from different codes were compared with each other, with the objective of reducing the uncertainties
% on such data, and in some sense, to define a \emph{calibration} procedure for a Solid-State Physics DFT code.
% Any other code able to extract the same quantities with the same approximations may be compared with the
% results of the community in order to assess its computational reliability.
% 
% For a theoretical approach that is based on numerical calculations, for which no analytic reference solution exists,
% reducing the computational uncertainty is the only possible way to shed light on the predictivity power of the
% model. In other terms, the ``accuracy'' of a result with respect to experimental data may be reliably quantified
% only when the computational uncertainty is guaranteed to be significantly lower than the found discrepancy.
% For communities like DFT, such a ``calibration'' of computer codes among each other is a discipline that is still in its infancy;
% the DeltaCodes initiative and the foundations of DFT are separated in time by more than 50 years, and the computational
% model employed in the comparison (the Perdew-Burke-Erzerhof functional~\cite{PBE}) has been introduced more than 20 years ago.

For a theoretical approach that is based on numerical calculations, for which no analytic reference solution exists,
reducing the computational uncertainty is the only possible way to shed light on the predictivity power of the
model. In other terms, the ``accuracy'' of a result with respect to experimental data may be reliably quantified
only when the computational uncertainty is guaranteed to be significantly lower than the found discrepancy.
For communities like DFT, the need of such a ``calibration'' of computer codes has led to the DeltaCodes project~\cite{deltaTest2016}. 
This project represents a remarkable example of precision-driven initiative where a grassroot community of code-owners struggled to extract 
well-defined computational results in the context of Density Functional Theory (DFT) calculations. Ground-State quantum mechanical quantities 
as lattice constants and bulk modulus of elementary crystals extracted from different codes were compared with each other, with the objective of reducing the uncertainties on such data, and in some sense, to define a \emph{calibration} procedure for a Solid-State Physics DFT code.
Any other code able to extract the same quantities with the same approximations may be compared with the
results of the community in order to assess its computational reliability.

@@

In this paper, we focus our attention to Linear-Response treatments of \emph{molecular} systems,
in particular referring to LR-TDDFT calculation. We will investigate the impact that
\emph{open} (i.e. Isolated) Boundary Conditions (BC) that have to be imposed to the system will have on the computational evaluation of LR quantities.
In order to do this, we perform analytic considerations on the equations beyond LR treatment.
\added[id=dal]{This analysis evidences that the description of molecular systems beyond the ground state unavoidably leads to the inclusion of delocalised
degrees of freedom above a specific energetic threshold.}
Such considerations will enable us to distinguish quantities that \textit{a priori} can be compared among different treatments - and therefore tested for reproducibility -
from others that would \emph{explicitly} depend of the employed computational treatment. \added[id=dal]{Moreover on the basis of this analysis we can assess
something about the reliability of the results provided by computational setups able to explicit include the presence of delocalised degrees of freedom compared
to the ones in which these degrees of freedom are excluded by design.}
\deleted[id=dal]{In particular, we will discuss the interplay between reproducible LR quantities and the \emph{locality} of the related objects.}

@@

\replaced[id=dal]{Let us start our discussion by reviewing some well-known concepts of quantum mechanic under a perspective that
will turn out to be very useful in the forthcoming considerations about linear-response quantities. In particular we revisit the impact 
that the Isolated BC have on the solutions of the time-independent Schr\"odinger Equation.}{Let us start our discussion by revisiting the 
impact that the Isolated BC have on the solutions of the time-independent Schr\"odinger Equation. We here present well-known concepts of 
quantum mechanics, under a perspective that will turn out to be very useful in the forthcoming considerations about linear-response quantities.}

@@

We can therefore introduce an explicit representation of the
response density parametrized as
\be\lb{rhoPrimeFluctuationStateDef1}
\dm'_\Phi(\omega) = \sum_p\left(\ketbra{\psi_p}{f_p^\Phi(-\omega)} + \ketbra{f_p^\Phi(\omega)}{\psi_p}\right) \;.
\ee
Here we have introduced the set of $\omega$-dependent \emph{fluctuation states}
$\ket{f_p^\Phi(\omega)} = \dm'_\Phi(\omega) \ket{\psi_p}$, that have values in the unoccupied subspace of $\hnot$.
\added[id=dal]{Fluctuation states are complex quantities and consequently $\dm'_\Phi(\omega)$ is not hermitian; nonetheless the $\omega$ dependence of \eqref{rhoPrimeFluctuationStateDef1} ensures that the hermitian nature of this operator is recovered in the time domain, as expected.}

An analysis of the Liouville equation \eqref{LiouvillianRhopomegaDef1} for the response density written in this fashion evidences that the equations of motion of fluctuation states (FS) are
written as a modified Sternheimer equation~\cite{mahan1980,giustino2012,giustino2014}
\be\lb{fluctuationStateEqMotion1}
\left[\omega - (\hnot-\eps_p)\right]\ket{f_p^\Phi(\omega)} = \op Q_0(\op\Phi(\omega)+\op V'[\dm'_\Phi](\omega))\ket{\psi_p} \;,
\ee
and we observe that the first-order Hamiltonian contains,
apart from the perturbing field, a further term $\op V'[\dm'_\Phi]$ due to the density-dependence of $\hnot$.
It is easy to see that the solution of equation \eqref{fluctuationStateEqMotion1} can be casted in the form of
a set of scattering problems:
\be\lb{rhoPrimeFluctuationStateDef2}
\ket{f_p^\Phi(\omega)} = \GH(\omega+\epsilon_p)\left(\op V\ket{f_p^\Phi(\omega)} + \ket{s_p^\Phi(\omega)} \right)\;,
\ee
with the state
$
\ket{s_p^\Phi(\omega)} = \op Q_0(\op \Phi(\omega) +\op V'[\dm'_\Phi](\omega) )\ket{\psi_p}
$ as a source term.

\replaced[id=dal]{Equation \eqref{rhoPrimeFluctuationStateDef2} is the convenient starting point to assess the localization properties of the FS as it is easy to see that, for each choice of $p$, the value $\omega = |\eps_p|$ represents the threshold level that governs the asymptotic behavior of the associated Helmholtz kernel.}{It can be seen that for each choice of $p$, the value $\omega = |\eps_p|$ represents the threshold level
that governs the localization properties of the FS.}
As a ulterior \emph{stringent} hypothesis, let us now assume that the source term $\ket{s_p^\Phi(\omega)}$ is \added[id=dal]{spatially} localized, \textit{i.e.} its evaluation can be restricted on a bounded real-space domain.
Even when such locality argument holds true, for $\omega > |\eps_p|$,
the Helmholtz kernel behaves as a spherical wave and  gives raise to a $\ket{f^\Phi_p(\omega)}$ which is \emph{delocalized} in the entire real-space domain.
Such delocalization is an intrinsic consequence of the value of $\omega$, and does not depend on the particular choice of
$\op \Phi$.
Only for $\omega < |\eps_p|$ the kernel contains an exponential damping factor and $\ket{f^\Phi_p(\omega)}$ exhibits a bound-state behavior in the long range,
once again provided that the above locality arguments on $\ket{s_p^\Phi(\omega)}$ -- hence on $\op V'$ and $\Phi$ -- are valid.

\added[id=dal]{Added a subsection for static perturbations}

\subsection{Localized behavior for static perturbations}

The locality property of the fluctuation state can be directly confirmed for static perturbations at $\omega=0$.
Let us consider a static perturbing operator $\op\Phi_s$ that can be restricted to a finite domain.
Any local operator, like \emph{e.g.} a static electric field $\mathbf F$, with associated scalar potential $\Phi_s[\mathbf F](\r) = -\mathbf F \cdot \r$, would satisfy such property.
We consider the perturbed Hamiltonian $\hnot[\dm_{\Phi_s}] + \op\Phi_s[\mathbf F]$, where $\dm_{\Phi_s}$ represents the GS density of the perturbed system, and assume that the
perturbation strength ($|\mathbf F|$ in the above example) is sufficiently small to be consistent with a LR treatment of the problem.
In this case, the occupied eigenfunction $\ket{\psi_p^{\Phi_s}}$ of the perturbed Hamiltonian can be extracted with traditional GS techniques,
as from the discussion around Eq.~\eqref{SEHelmholtz1}, they exhibit bound-state behavior.
\begin{figure}[t]
\includegraphics[scale=0.68]{Fig1_CO_statPolvsBox.eps}
\caption{\label{co_alphaStatic}(Color online) Convergence of the statical polarizability $\alpha_{zz}$ (blue continuous line) and of the ground state dipole $d_z$ (red dotted line)
of $CO$ molecule as a function of the size of the computational domain. The inset evidences the relative convergence (in percent) of the two quantities w.r.t. the corresponding value associated to the largest
box.}
\end{figure}

@@

\added[id=dal]{Modified the name of the subsection}

\subsection{Localization propeties at finite $\omega$. Emergence of a strong and a weak regime}

%\subsection{Strong and weak convergence at finite $\omega$}

@@

\added[id=dal]{It is possible to reformulate this type of analysis by directly looking at LR quantities that are defined independently from the perturbing operators and target observables. This can be achieved by studying the
\emph{susceptibility functional}, defined implicitly trough the operator equation}
\deleted[id=dal]{
In view of reproducibility, the scalar product of Eq.~\eqref{LinearResponseFunctDef1}, defining the observable fluctuation, may be compared among various numerical treatments.
However, even though directly related to experimental results, such quantity is not by itself a property of the molecule, as it is a function of $\omega$ that explicitly depends
on the perturbation operator $\op \Phi$, not to mention the target observable $\op O$. Alternatively, we may consider only the fluctuation state, but it is still a $\op \Phi$-dependent wavefunction,
with a potential local behavior only for $\omega$ below the ionization threshold. Hence, comparing fluctuation states among diverse numerical treatments would be a
challenging task.}
\deleted[id=dal]{It is therefore much more interesting to concentrate on LR quantities that are defined independently from the perturbing operators and target observables.
This can be achieved by studying the
\emph{susceptibility functional}, defined implicitly trough the operator equation}
\be\lb{SusceptibilityDef1}
\dm'(\omega) = \left(\omega -\mathcal L \right)^{-1} \commutator{\op \Phi(\omega)}{\dmnot} = \sop{{\cal \chi}}(\omega) \op\Phi(\omega) \;.
\ee
The susceptibility (super)operator $\cal \chi(\omega)$ is therefore closely related to the resolvent of the Liouvillian of Eq.~\eqref{LiouvillianRhopomegaDef1}, which is completely defined
from the molecule Hamiltonian $\hnot[\dm]$.
The ``excitation modes''  of the molecule are defined through the \emph{excitation operators}, satisfying
\be\lb{ExcitationOperatorsDef1}
\Liouv \op E_a = \Omega_a \op E_a \qq \op{\tilde E}_a \Liouv = \Omega_a \op{\tilde E}_a \;,
\ee
together with the operator orthonormalization condition %$\trace{\op{\tilde E}\op E'} = \delta(E-E')$.
\be\lb{orthoExcitatioOpDef1}
\trace{\op{\tilde E}_a\op E_b} = \delta_{ab} \;.
\ee
Excitations defined in that way may be considered as a basis of the operators that satisfy the transverse condition as in
Eq.~\eqref{RhopTransverseDef1}, and enable us to express the susceptibility as a spectral decomposition:
\be\lb{RhopExcitationDef1}
\sop{{\cal \chi}}(\omega) \cdot   = \sum_{\{a\}} \op E_{a} \,
\frac{\trace{\commutator{\op{\tilde E}_a}{\dmnot}\cdot}}{\omega-\Omega_a} \;.
\ee
\added[id=dal]{Here add a short justification for the expression of the real and imaginary part of the linear response functional written
in terms of excitation operators. Principal part vs delta functions.}

\added[id=dal]{Modified name of the subsection}

\subsection{Localization features of the excitation operators. The analytic structure of linear susceptibility}

Each excitation ``mode'' of the system $\op E_a$ with energy $\Omega_a$, can be associated to a set of states $\{\phi^a_p,\chi^a_p\}_{p \in a}$,
entirely defined in the unoccupied subspace.
In this way we can parametrize, without loss of generality, the excitation operators of the system as:
\begin{align}\lb{ExcitationOperatorsDef2}
\op E_a &= \sum_p \left( \ket{\phi^a_p}\bra{\psi_p} + \ket{\psi_p} \bra{\chi^a_p}\right)\;, \nn \\
\op{\tilde E}_a = \commutator{\dmnot}{\op E^t_a} &= \sum_p \left( \ket{\psi_p}\bra{\phi^a_p} - \ket{\chi^a_p} \bra{\psi_p}\right)\;.
\end{align}
For each choice of $\{\phi^a_p,\chi^a_p\}_{p \in a}$, these objects represent, respectively, the state in which $\ket{\psi_p}$
is excited -- or from whom it decays -- when the system is subject to the \emph{monochromatic} perturbation $\op \Phi_a \equiv \commutator{\dmnot}{\op E_a}$.
We denote with the shorthand $p \in a$ the fact that the sum of Eq.~\eqref{ExcitationOperatorsDef2}
is in general restricted only to a $a$-dependent subset of the occupied orbitals, depending on the particular symmetry
of the excitation.
Excited states  satisfy the normalization condition
\be\lb{ExcitedStateOrthNormDef1}
\sum_p \left(\brket{\phi_p^a}{\phi_p^b} - \brket{\chi_p^b}{\chi_p^a}\right) = \delta_{ab} \;.
\ee
By applying \eqref{ExcitationOperatorsDef1} to the above parametrization
of the excitations, we obtain the following Sternheimer-like equations for the excited states:
\begin{align}\lb{ExcitationOperatorsDef3}
&\left[\Omega_a-(\hnot - \eps_p)\right] \ket{\phi_p^a} = \op Q_0 \op V'[\op E_a] \ket{\psi_p} \;, \nn\\
&\bra{\chi_p^a}\left[-\Omega_a-(\hnot - \eps_p)\right] = \bra{\psi_p} \op V'[\op E_a] \op Q_0  \;.
\end{align}
By writing these equations in a basis set of virtual states we obtain the well-known Casida's eigenvalue equation\cite{CasidaBook}
associated to the excitation energy $\Omega_a$ (see appendix \ref{casida}).
The spectrum is symmetric w.r.t the inversion of the eigenvalues
$\Omega_a \rightarrow -\Omega_a$ and the associated excitation with negative energy are described by the transposed pair $\{\chi^a_p,\phi^a_p\}$.
We concentrate our analysis on the sector of positive energies, i.e. $\Omega_a > 0$.
